---
title: "Practice assignment 7"
author: "e2u"
date: "15/12/2021"
output: html_document
---


```{r}
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(vioplot)

```
```{r}
titanic <- read.csv("titanic.csv")
head(titanic)
```

```{r}
#NA's, missing values and summary

colSums(is.na(titanic))
summary(titanic)
colSums(titanic=="")
```
```{r}
#Bar plot of gender
ggplot(titanic) + geom_bar(aes(x = Sex, fill = Sex))
```
Bar plot of ticket class
```{r}
#converts Pclass to categorical variable
titanic$Pclass <- as.factor(titanic$Pclass)
head(titanic)
```

```{r}
#Bar plot of ticket class 
ggplot(titanic) + geom_bar(aes(x = Pclass, fill = Pclass))

```

Bar plot of Survival 
```{r}
#converts Survival to categorical variable
titanic$Survived <- as.factor(titanic$Survived)

```

```{r}
#Bar plot of Survived
ggplot(titanic) + geom_bar(aes(x = Survived, fill = Survived))

```

¤ Question 2:
```{r}
#Histogram for passenger age with minimal theme

ggplot(titanic, aes(x = Age, fill = Sex)) + 
    geom_histogram()+
    labs(title = "Histogram of passenger age", x = "Age", y = "Count") +
    theme_minimal()
```

```{r}
#density plot passenger age with minimal theme

ggplot(titanic, aes(x = Age, fill = Sex)) + 
    geom_density(alpha = .8) +
    labs(title = "Density plot of passenger age", x = "Age", y = "Count") +
    theme_minimal()
```
Describe the passengers' age with boxplots:

```{r}
#Boxplot for age per ticket class
ggplot(titanic) + geom_boxplot(aes(x = Age, fill = Pclass)) +
    labs(title = "Box plot of age per ticket class", x = "Age")
```
```{r}
#Boxplot for age based on survival
ggplot(titanic) + geom_boxplot(aes(x = Age, fill = Survived)) +
    labs(title = "Box plot of age based on survival", x = "Age")
```
We could explore the data further for age ranges:
```{r}
#Create age categories
categories<-c('0-10','10-20','20-30','30-40','40-50','50-60','60-70','70-80')
titanic$Age_cat<-cut(titanic$Age, seq(0,80,10),labels=categories)
Age_count<-table(titanic$Age_cat)
Age_count
```
```{r}
# we don't the exclude NA's as they represent ticket class of passengers without age data.

ggplot(titanic) + geom_bar(position = "dodge") + 
  aes(x = Age_cat, fill = Pclass) +
  labs(title = "Bar plot of age per ticket class ", x = "Age", y = "Count")
```
Age based on survival:
```{r}
ggplot(titanic) + geom_bar(position = "dodge") + 
    aes(x = Age_cat, fill = Survived)+
    labs(title = "Bar plot of age per ticket class ", x = "Age", y = "Count")
```

Histogram for the travel fare and a table showing the number of
people who did not pay
```{r}
ggplot(titanic, aes(x = Fare)) + 
    geom_histogram()+
    labs(title = "Histogram of Fare", x = "Price", y = "Count")
```
The histogram shows an extreme outlier, it seems like the passenger who paid 512.33(Lesurer, Mr. Gustave J) is either a mistake or have rented an very exclusive suite.
A Google search revealed that Mr. Gustave J. lesueur was a servant for Thomas Cardeze, the suite was occupied by Thomas, his mother and two servants. 
ref: [https://medium.com/silk-stories/women-and-children-first-9273e97289b0]



Histogram and table showing no. of people who didn't pay

```{r}
#Add table of non paying travelers to the histogram
Pas.count <- sum(titanic$Fare == "0")
h2 <- cbind(c("Non paying passengers:"), Pas.count) 

#Histogram
ggplot(titanic, aes(x = Fare, color = Fare)) + 
    geom_histogram(binwidth = 10)+
    labs(title = "Histogram of Fare", x = "Price", y = "Count") + 
    annotation_custom(tableGrob(h2), xmin=200, xmax=600, ymin=350, ymax=400)
```


A chart to describe the family size per ticket class

```{r}
titanic$Family_class <- titanic$SibSp + titanic$Parch+1
ggplot(data = titanic[!is.na(titanic$Family_class),],aes(x=Family_class, fill=Pclass))+geom_bar(position = "dodge")+ scale_x_continuous(limits = c(2, 12), breaks = 2:12) + 
labs(title = "Histogram of family size per ticket class", x = "Size", y = "Count")
```
We could also describe the family size and survival rate

```{r}
titanic$Family_class <- titanic$SibSp + titanic$Parch+1
ggplot(data = titanic[!is.na(titanic$Family_class),],aes(x=Family_class, fill=Survived))+geom_bar(position = "dodge")+ scale_x_continuous(limits = c(2, 12), breaks = 2:12) + 
labs(title = "Histogram of family size per survival", x = "Size", y = "Count")
```

Series of stacked bar charts to show the how survival differs for different
gender and ticket class

```{r}
# with faceting  
ggplot(titanic, aes(x = Pclass, fill = Survived)) +   
    geom_bar() +   
    facet_grid(. ~ Sex) +  
    xlab("Passenger Class") +
    ylab("Count") + 
    scale_fill_manual("Survived", values = alpha( c("#F8766D", "#00BFC4")))
    
```

It looks like that females on 1st and 2nd class had an good probability of survival, to demonstrate this we can divide Passenger class by gender to see the survival percentage:

```{r}
ggplot(titanic,aes(x=Pclass, fill=Survived))+geom_bar(position="fill")+
  facet_wrap(~Sex)+xlab("Passengers Class")+
  ylab("Percentage")
```

Violin chart describing how survival relates to age and gender
```{r}
ggplot(titanic, aes(x=Sex, y=Age, fill=Survived)) +
  geom_violin()
```
Violin chart describing the survival rate related to age and ticket class
```{r}
ggplot(titanic, aes(x=Pclass, y=Age, fill=Survived)) +
  geom_violin()
```


- Generally passengers on 1st and 2nd class were more likely to survive, indicating that the security might have been less on 3rd. class
- Females were more likely than males to survive in each class. 
- Males on 3rd. class was unlikely to survive.
- Large family's were less likely to survive.
- Passengers of 40+ age are more likely to survive on 1st. class than 3rd.
